---
Author: Eniola Oladimeji
title: "Languages in Canada"
date: 2026-09-22
---

Data: [can_lang](https://github.com/ttimbers/canlang) from the canlang package by Tiffany Timbers (MIT licence). Source: Statistics Canada, Census of Population, 2016. Reproduced and distributed on an "as is" basis with the permission of Statistics Canada.

The goal of this analysis is to explore how Aboriginal languages are used in Canada: how many people speak them at work relative to how many have them as a mother tongue, and which languages are being learned as second languages.


```{r}
#| message: false
#| warning: false
library(tidyverse)

can_lang <- read_csv("can_lang.csv")
head(can_lang)
```

```{r}
aboriginal <- can_lang |>
  filter(category == "Aboriginal languages")

aboriginal
```

@fig-top-aboriginal shows the ten Aboriginal languages with the most mother-tongue speakers.

```{r}
#| label: fig-top-aboriginal
#| fig-cap: "Top 10 Aboriginal languages in Canada by number of mother-tongue speakers (2016 Census)."
aboriginal |>
  slice_max(mother_tongue, n = 10) |>
  ggplot(aes(x = mother_tongue, y = reorder(language, mother_tongue))) +
  geom_col() +
  labs(x = "Mother-tongue speakers", y = "Language")
```

```{r}
aboriginal |>
  mutate(pct_at_work = most_at_work / mother_tongue * 100) |>
  select(language, mother_tongue, most_at_work, pct_at_work) |>
  arrange(desc(pct_at_work))
```

```{r}
aboriginal |>
  mutate(second_lang = lang_known - mother_tongue) |>
  select(language, second_lang) |>
  arrange(desc(second_lang))
```

Steps I took:
First, I read in the data with `readr` package from tidyverse. Then I filtered the data for aboriginal languages spoken in Canada. 
Next, I found the percentage of people who have Aboriginal languages as their mother tongue and speak it at work. Lastly and just for curiosity, found the aboriginal languages being learned as a second language. 
